Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-07-01 manuscript.Rnw
Minor edits based on NCC reviews.

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation\cite{ise2010}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass determines turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \cite{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass\citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014}.

@@

Here, we use a unique, new database\citep{Falster2015} (Fig. 1) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. The dataset includes over 21,000 observations of biomass and size metrics for individual plants, contributed by 175 studies across nine vegetation types (Fig. S1). Previous work based on whole stands\citep{oneilldeangelis1981,enquistniklas2002,reich2014} and very small plants\citep{poorter2012}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). Here we extend this finding to show that LMF differs not only across the angiosperm-gymnosperm divide, but also within angiosperms, at the level of individual plants, and across the entire size range of woody plants in natural forests (0.1-100m) (Fig. 2a).

Moreover, we show that differences in LMF among woody PFTs arise due to differences in leaf morphology, but they converge in the amount of leaf area held on the plant (Fig. 2b-c). Previously, it was unclear whether PFTs with higher leaf biomass also had higher total leaf area. Leaf area is relevant because it drives total light interception and thus photosynthesis. Several classic studies have assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}. We found no evidence for this, instead we found that the three PFTs maintain similar leaf area per unit above-ground biomass (the `leaf area ratio`, LAR).

@@

We only used LMA directly estimated for the harvested plants (typically for a subsample of leaves). For conifers, leaf area was converted to half-total surface area using published conversion factors \citep{barclay2000}, which is most appropriate for comparison to flat leaves \citep{lang1991, chen1992}. Stem cross-sectional area was measured either at breast height (ca. 1.3m) and/or at the base of the plant. For the subset of the data where both were measured, we estimated basal stem area ($A_S$) from breast height stem area ($A_{Sbh}$) using a non-linear regression model (see~\ref{sec:predas}).

@@

Figure 1a uses an image by Lana Heydon, IAN Image Library http://ian.umces.edu/imagelibrary. We thank Martin de Kauwe for discussions and comments on an earlier version of this paper.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure2.pdf}
    \caption{\textbf{Dominant woody PFTs differ in leaf mass fraction due to underlying differences in leaf construction cost. } (a) Leaf mass fraction (LMF = leaf mass / above-ground biomass) by PFT. Each symbol is an individual plant. Lines are generalized additive model fits. (b) and (c) Leaf mass fraction (LMF) and leaf area ratio (LAR) at the average plant height in the dataset, estimated from panel (a). (d) Average leaf mass per unit basal stem area, and (e) leaf area per unit basal stem area for the three major PFTs confirm that the between-PFT variation in leaf mass fraction is due to leaf mass per unit basal area. Error bars are 95\% confidence intervals. Letters denote significant differences (at $\alpha = 0.05$).}
    \label{fig:figure2}
\end{figure}

>>>>

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation\cite{ise2010}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \cite{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass\citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014}.

@@

Here, we use a unique, new database\citep{Falster2015} (Fig. 1) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. The dataset includes over 21,000 observations of biomass and size metrics for individual plants, contributed by 175 studies across nine vegetation types (Fig. S1). Previous work based on whole stands\citep{oneilldeangelis1981,enquistniklas2002,reich2014} and individual plants\citep{poorter2012}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). Here we extend this finding to show that LMF differs not only across the angiosperm-gymnosperm divide, but also within angiosperms, at the level of individual plants, and across the entire size range of woody plants in natural forests (0.1-100m) (Fig. 2a).

Moreover, we show that differences in LMF among woody PFTs arise due to differences in leaf morphology, but they converge in the amount of leaf area held on the plant (Fig. 2b-c). Previously, it was unclear whether PFTs with higher leaf biomass also had higher total leaf area. Leaf area is relevant because it drives total light interception and thus photosynthesis. Several classic studies have assumed, based only on differences in measurements of leaf mass, that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}. We found no evidence for this, instead we found that the three PFTs maintain similar leaf area per unit above-ground biomass (the `leaf area ratio`, LAR).

@@

We only used LMA directly estimated for the harvested plants (typically for a subsample of leaves). For conifers, leaf area was converted to half-total surface area using the average of a set of published conversion factors \citep{barclay2000}. We used a different conversion for pines (\emph{Pinus} spp.) than non-pines, because the conversion factors were very different for these two groups \citep{barclay2000}. This conversion was necessary because half-total surface area is most appropriate for comparison to flat leaves \citep{lang1991, chen1992}. Stem cross-sectional area was measured either at breast height (ca. 1.3m) and/or at the base of the plant. For the subset of the data where both were measured, we estimated basal stem area ($A_S$) from breast height stem area ($A_{Sbh}$) using a non-linear regression model (see~\ref{sec:predas}).

@@

Figure 1a uses an image by Lana Heydon, IAN Image Library http://ian.umces.edu/imagelibrary. Our sincere thanks to everyone who contributed data to the Biomass and Allometry Database. We thank Martin de Kauwe for discussions and comments on an earlier version of this paper.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure2.pdf}
    \caption{\textbf{Dominant woody PFTs differ in leaf mass fraction due to underlying differences in leaf construction cost. } (a) Leaf mass fraction ($M_F/M_T$ = leaf mass / above-ground biomass) by PFT. Each symbol is an individual plant. Lines are generalized additive model fits. (b) and (c) Leaf mass fraction and leaf area ratio ($A_F/M_T$) at the average plant height in the dataset, estimated from panel (a). (d) Average leaf mass per unit basal stem area, and (e) leaf area per unit basal stem area for the three major PFTs confirm that the between-PFT variation in leaf mass fraction is due to leaf mass per unit basal area. Error bars are 95\% confidence intervals. Letters denote significant differences (at $\alpha = 0.05$).}
    \label{fig:figure2}
\end{figure}
